---
title: "Plastic MFA results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("C:/Users/mingxu/Box Sync/Paper/Plastics MFA/mingxu-um.github.io/plastics.RData")
```

## 1. Consumption vs. population

## 1.1 Total consumption vs. population 1950-2017
```{r pressure, echo=FALSE}
#plots total consumption vs. population
n_country=dim(consumption)[3]
n_year=dim(consumption)[1]
color=rainbow(n_country)
plot(socioeconomic[,1,1],consumption[,7,1],xlim=c(100,max(socioeconomic[,1,])),ylim=c(0.01,max(consumption[,7,])),col=color[1],log="xy",xlab="population (1,000)",ylab="total plastic consumption (kton)",type = "l")
for (i in c(2:n_country)){
  points(socioeconomic[,1,i],consumption[,7,i],col=color[i],type="l")
  text(socioeconomic[n_year,1,i],consumption[n_year,7,i],dimnames(consumption)[[3]][i])
}
```

### 1.2 It looks like the plastic consumption of all countries grows with popluation exponentially (linearly on a log-log scale), except every country started differently in 1950. Here is a plot of plastic consumption of countries in 1950 vs. population:

```{r, echo=FALSE, warning=FALSE}
plot(socioeconomic[1,1,],consumption[1,7,],xlim=c(100,max(socioeconomic[,1,])),ylim=c(0.01,max(consumption[,7,])),col=color[1],log="xy",xlab="population-1950 (1,000)",ylab="total plastic consumption-1950 (kt)")
```

### log-log regression of total plastic consumption vs. population in 1950:
```{r, echo=FALSE}
lm_log.model=lm(log1p(consumption[1,7,])~log1p(socioeconomic[1,1,]))
summary(lm_log.model)
```

### consumption vs. pop normalized by 1950
```{r, echo=FALSE}
normalized_pop=t(t(socioeconomic[,1,])/socioeconomic[1,1,])
nor_consumption = t(t(consumption[,7,])/consumption[1,7,])
plot(1,1,xlim=c(0,5),ylim=c(0,4000),xlab="population (1950=1)",ylab="total plastic consumption (kt) (1950=1)")
for (i in c(1:n_country)){
  points(normalized_pop[,i],nor_consumption[,i],col=color[i],type="l")
  text(normalized_pop[n_year,i],nor_consumption[n_year,i],dimnames(consumption)[[3]][i])
}
```

### If logY = a*logP + b, and logY0 = m*logP0 + n, we can get logYn = a*logPn + (a-m)*logP0 + (b-n). The results are not good, probably because these countries are two clusters (see above figure)
```{r, echo=FALSE}
logYn = log1p(nor_consumption)
logPn = log1p(normalized_pop)
logP0 = matrix(1,68,42) %*% diag(log1p(socioeconomic[1,1,]))
logYn[which(is.nan(logYn))] = NA
logYn[which(logYn==Inf)] = NA
logYn=as.vector(logYn[,-c(20,23,25,37,38)])
logPn=as.vector(logPn[,-c(20,23,25,37,38)])
logP0=as.vector(logP0[,-c(20,23,25,37,38)])
lm_model.model=lm(logYn~logPn+logP0, na.action = na.exclude)
summary(lm_model.model)
```

### Range between (0,1) after normalization
```{r, echo=FALSE}
stdz_pop=normalized_pop
stdz_consumption=nor_consumption
for (i in c(1:n_country)){
  max_p=max(stdz_pop[,i])
  min_p=min(stdz_pop[,i])
  stdz_pop[,i]=(stdz_pop[,i]-min_p)/(max_p-min_p)
  max_c=max(stdz_consumption[,i])
  min_c=min(stdz_consumption[,i])
  stdz_consumption[,i]=(stdz_consumption[,i]-min_c)/(max_c-min_c)
}
plot(0,0,xlim=c(0,1),ylim=c(0,1),xlab="normalized pop (0,1)",ylab="normalized consumption (0,1)")
for (i in c(1:n_country)){
  points(stdz_pop[,i],stdz_consumption[,i],col=color[i],type="l")
  text(stdz_pop[n_year,i],stdz_consumption[n_year,i],dimnames(consumption)[[3]][i])
}

#log transformation first then normalization
logCspt=log10(consumption[,7,])
logPop=log10(socioeconomic[,1,])
for (i in c(1:n_country)){
  max_c=max(logCspt[,i])
  min_c=min(logCspt[,i])
  logCspt[,i]=(logCspt[,i]-min_c)/(max_c-min_c)
  max_p=max(logPop[,i])
  min_p=min(logPop[,i])
  logPop[,i]=(logPop[,i]-min_p)/(max_p-min_p)
}
plot(0,0,xlim=c(0,1),ylim=c(0,1),xlab="normalized log_pop (0,1)",ylab="normalized log_consumption (0,1)")
for (i in c(1:n_country)){
  points(logPop[,i],logCspt[,i],col=color[i],type="l")
  text(logPop[n_year,i],logCspt[n_year,i],dimnames(consumption)[[3]][i])
}
```

### as shown above, normalization after log transformation seems to be able to get a good fit


### 1.3 Now total plastic consumption/person vs. population in 1950:
```{r, echo=FALSE, warning=FALSE}
plot(socioeconomic[1,1,],consumption[1,7,]/socioeconomic[1,1,],log="xy", col=color[1],xlab="population-1950 (1,000)",ylab="total plastic consumption/capita-1950 (t/capita)")
for (i in c(1:n_country)){
  text(socioeconomic[1,1,i],consumption[1,7,i]/socioeconomic[1,1,i],dimnames(consumption)[[3]][i])
}
boxplot(consumption[1,7,]/socioeconomic[1,1,],ylab="total plastic consumption/person-1950 (t/capita)")
summary(consumption[1,7,]/socioeconomic[1,1,])
```

### 1.4 Finally plastic consumption/person vs. population from 1950-2017:
```{r, echo=FALSE, warning=FALSE}
plot(socioeconomic[,1,1],consumption[,7,1]/socioeconomic[,1,1],xlim=c(100,max(socioeconomic[,1,])),ylim=c(0.00001,max(consumption[,7,]/socioeconomic[,1,])),col=color[1],log="xy",xlab="population (1,000)",ylab="total plastic consumption/capita (t/capita)",type = "l")
for (i in c(2:n_country)){
  points(socioeconomic[,1,i],consumption[,7,i]/socioeconomic[,1,i],col=color[i],type="l")
  text(socioeconomic[n_year,1,i],consumption[n_year,7,i]/socioeconomic[n_year,1,i],dimnames(consumption)[[3]][i])
}
boxplot(as.vector(consumption[,7,]/socioeconomic[,1,]),ylab="total plastic consumption/capita (t/capita)")
summary(as.vector(consumption[,7,]/socioeconomic[,1,]))
```

### 1.5 By sectors (plastic consumption (kt) vs. population (1,000))

```{r, echo=FALSE, warning=FALSE}
plot(socioeconomic[,1,1],consumption[,7,1],xlim=c(100,max(socioeconomic[,1,])),ylim=c(0.001,max(consumption[,7,])),col=color[1],log="xy",type = "l",xlab="", ylab="",main="total")
for (i in c(2:n_country)){
  points(socioeconomic[,1,i],consumption[,7,i],col=color[i],type="l")
}
plot(socioeconomic[,1,1],consumption[,1,1],xlim=c(100,max(socioeconomic[,1,])),ylim=c(0.001,max(consumption[,1,])),col=color[1],log="xy",type = "l",xlab="", ylab="",main="Packaging")
for (i in c(2:n_country)){
  points(socioeconomic[,1,i],consumption[,1,i],col=color[i],type="l")
}
plot(socioeconomic[,1,1],consumption[,2,1],xlim=c(100,max(socioeconomic[,1,])),ylim=c(0.001,max(consumption[,2,])),col=color[1],log="xy",type = "l",xlab="", ylab="",main="Building/Construction")
for (i in c(2:n_country)){
  points(socioeconomic[,1,i],consumption[,2,i],col=color[i],type="l")
}
plot(socioeconomic[,1,1],consumption[,3,1],xlim=c(100,max(socioeconomic[,1,])),ylim=c(0.001,max(consumption[,3,])),col=color[1],log="xy",type = "l",xlab="", ylab="",main="Transportation")
for (i in c(2:n_country)){
  points(socioeconomic[,1,i],consumption[,3,i],col=color[i],type="l")
}
plot(socioeconomic[,1,1],consumption[,4,1],xlim=c(100,max(socioeconomic[,1,])),ylim=c(0.001,max(consumption[,4,])),col=color[1],log="xy",type = "l",xlab="", ylab="",main="EE")
for (i in c(2:n_country)){
  points(socioeconomic[,1,i],consumption[,4,i],col=color[i],type="l")
}
plot(socioeconomic[,1,1],consumption[,5,1],xlim=c(100,max(socioeconomic[,1,])),ylim=c(0.001,max(consumption[,5,])),col=color[1],log="xy",type = "l",xlab="", ylab="",main="HH")
for (i in c(2:n_country)){
  points(socioeconomic[,1,i],consumption[,5,i],col=color[i],type="l")
}
plot(socioeconomic[,1,1],consumption[,6,1],xlim=c(100,max(socioeconomic[,1,])),ylim=c(0.001,max(consumption[,6,])),col=color[1],log="xy",type = "l",xlab="", ylab="",main="Other")
for (i in c(2:n_country)){
  points(socioeconomic[,1,i],consumption[,6,i],col=color[i],type="l")
}
```

## 2. Over time

## 2.1 Consumption over time

```{r, echo=FALSE}
for (j in c(1:dim(consumption)[2])){
  plot(0,0,xlim=c(1950,2017),ylim=c(0,1),type="l",main=dimnames(consumption)[[2]][j])
  for (i in c(1:n_country)){
    points(c(1950:2017),consumption[,j,i]/max(consumption[,j,i]),col=color[i],type="l")
  }
}
```

## 2.2 Waste over time

```{r, echo=FALSE}
for (j in c(1:dim(waste)[2])){
  plot(0,0,xlim=c(1950,2017),ylim=c(0,1),type="l",main=dimnames(waste)[[2]][j])
  for (i in c(1:n_country)){
    points(c(1950:2017),waste[,j,i]/max(waste[,j,i]),col=color[i],type="l")
  }
}
```

## 2.3 Stock over time
```{r, echo=FALSE}
for (j in c(1:dim(stock)[2])){
  plot(0,0,xlim=c(1950,2017),ylim=c(0,1),type="l",main=dimnames(stock)[[2]][j])
  for (i in c(1:n_country)){
    points(c(1950:2017),stock[,j,i]/max(stock[,j,i]),col=color[i],type="l")
  }
}
```

## 3. Consumption vs. GDP
nothing interesting

```{r, echo=FALSE, warning=FALSE}
# total vs. GDP
par(mfrow=c(1,2))
plot(0,0,log="xy",xlim=c(0.1,max(socioeconomic[,2,])),ylim=c(0.001,max(consumption[,7,])),type = "l",xlab="GDP current ($)", ylab="consumption total (kton)",main="Total")
for (i in c(1:n_country)){
  points(socioeconomic[,2,i],consumption[,7,i],col=color[i],type="l")
}
plot(0,0,log="xy",xlim=c(0.1,max(socioeconomic[,3,])),ylim=c(0.001,max(consumption[,7,])),type = "l",xlab="GDP constant ($)", ylab="consumption total (kton)",main="Total")
for (i in c(1:n_country)){
  points(socioeconomic[,3,i],consumption[,7,i],col=color[i],type="l")
}

# total/capita vs. GDP
par(mfrow=c(1,2))
plot(0,0,log="xy",xlim=c(0.1,max(socioeconomic[,2,])),ylim=c(0.00001,max(consumption[,7,]/socioeconomic[,1,])),type = "l",xlab="GDP current ($)", ylab="consumption/capita (t/a)",main="Total/capita")
for (i in c(1:n_country)){
  points(socioeconomic[,2,i],consumption[,7,i]/socioeconomic[,1,i],col=color[i],type="l")
}
plot(0,0,log="xy",xlim=c(0.1,max(socioeconomic[,3,])),ylim=c(0.00001,max(consumption[,7,]/socioeconomic[,1,])),type = "l",xlab="GDP constant ($)", ylab="consumption/capita (t/a)",main="Total/capita")
for (i in c(1:n_country)){
  points(socioeconomic[,3,i],consumption[,7,i]/socioeconomic[,1,i],col=color[i],type="l")
}

# total vs. GDP/capita
par(mfrow=c(1,2))
plot(0,0,log="xy",xlim=c(0.1,max(socioeconomic[,4,])),ylim=c(0.001,max(consumption[,7,])),type = "l",xlab="GDP current/person ($/1,000)", ylab="consumption total (kton)",main="Total")
for (i in c(1:n_country)){
  points(socioeconomic[,4,i],consumption[,7,i],col=color[i],type="l")
}
plot(0,0,log="xy",xlim=c(0.1,max(socioeconomic[,5,])),ylim=c(0.001,max(consumption[,7,])),type = "l",xlab="GDP constant/person ($/1,000)", ylab="consumption total (kton)",main="Total")
for (i in c(1:n_country)){
  points(socioeconomic[,5,i],consumption[,7,i],col=color[i],type="l")
}

# total/capita vs. GDP/capita
par(mfrow=c(1,2))
plot(0,0,log="xy",xlim=c(0.1,max(socioeconomic[,4,])),ylim=c(0.00001,max(consumption[,7,]/socioeconomic[,1,])),type = "l",xlab="GDP current/person ($/1,000)", ylab="consumption/capita (t/a)",main="Total/capita")
for (i in c(1:n_country)){
  points(socioeconomic[,4,i],consumption[,7,i]/socioeconomic[,1,i],col=color[i],type="l")
}
plot(0,0,log="xy",xlim=c(0.1,max(socioeconomic[,5,])),ylim=c(0.00001,max(consumption[,7,]/socioeconomic[,1,])),type = "l",xlab="GDP constant/person ($/1,000)", ylab="consumption/capita (t/a)",main="Total/capita")
for (i in c(1:n_country)){
  points(socioeconomic[,5,i],consumption[,7,i]/socioeconomic[,1,i],col=color[i],type="l")
}
```

## 4. consumption vs. urbanization
```{r, echo=FALSE, warning=FALSE}
# consumption vs urbanization
par(mfrow=c(1,2))
plot(0,0,log="y",xlim=c(0,1),ylim=c(0.00001,max(consumption[,7,])),type = "l",xlab="urbanization rate", ylab="consumption (kt)",main="Total vs. urbanization")
for (i in c(1:n_country)){
  points(socioeconomic[,6,i],consumption[,7,i],col=color[i],type="l")
}
plot(0,0,log="xy",xlim=c(1,max(socioeconomic[,7,])),ylim=c(0.00001,max(consumption[,7,])),type = "l",xlab="urban pop (1,000)", ylab="consumption (kt)",main="Total vs. urban pop")
for (i in c(1:n_country)){
  points(socioeconomic[,7,i],consumption[,7,i],col=color[i],type="l")
}

# consumption/capita vs. urbanization
par(mfrow=c(1,2))
plot(0,0,log="y",xlim=c(0,1),ylim=c(0.00001,max(consumption[,7,]/socioeconomic[,1,])),type = "l",xlab="urbanization rate", ylab="consumption/capita (t/c)",main="Total/capita vs. urbanization")
for (i in c(1:n_country)){
  points(socioeconomic[,6,i],consumption[,7,i]/socioeconomic[,1,i],col=color[i],type="l")
}
plot(0,0,log="xy",xlim=c(1,max(socioeconomic[,7,])),ylim=c(0.00001,max(consumption[,7,]/socioeconomic[,1,])),type = "l",xlab="urban pop (1,000)", ylab="consumption/capita (t/c)",main="Total/capita vs. urban pop")
for (i in c(1:n_country)){
  points(socioeconomic[,7,i],consumption[,7,i]/socioeconomic[,1,i],col=color[i],type="l")
}

# consumption/urban pop vs. urbanization
par(mfrow=c(1,2))
plot(0,0,log="y",xlim=c(0,1),ylim=c(0.00001,max(consumption[,7,]/socioeconomic[,7,])),type = "l",xlab="urbanization rate", ylab="consumption/urban pop (t/c)",main="Total/urban pop vs. urbanization")
for (i in c(1:n_country)){
  points(socioeconomic[,6,i],consumption[,7,i]/socioeconomic[,7,i],col=color[i],type="l")
}
plot(0,0,log="xy",xlim=c(1,max(socioeconomic[,7,])),ylim=c(0.00001,max(consumption[,7,]/socioeconomic[,7,])),type = "l",xlab="urban pop (1,000)", ylab="consumption/urban pop (t/c)",main="Total/uban pop vs. urban pop")
for (i in c(1:n_country)){
  points(socioeconomic[,7,i],consumption[,7,i]/socioeconomic[,7,i],col=color[i],type="l")
}
```